---
title: "Evaluating Content-Related Validity Evidence Using Text Modeling"
shorttitle: "Text Modeling Validity"
date: "`r format(Sys.time(), '%B %d, %Y')`"

author: 
  - name          : "Daniel Anderson"
    affiliation   : "1"
    corresponding : yes
    email         : "daniela@uoregon.edu"
    address       : "5262 University of Oregon"
  - name          : "Brock Rowley"
    affiliation   : "1"
    email         : "brockr@uoregon.edu"
  - name          : "Sondra Stegenga"
    affiliation   : "1"
    email         : "sondras@uoregon.edu"
  - name          : "P. Shawn Irvin"
    affiliation   : "1"
    email         : "pirvin@uoregon.edu"
  - name          : "Joshua M. Rosenberg"
    affiliation   : "2"
    email         : "jmrosenberg@utk.edu"
affiliation:
  - id            : "1"
    institution   : "University of Oregon"
  - id            : "2"
    institution   : "University of Tennessee"

bibliography      : refs.bib

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no
mask              : no

abstract: |
The alignment of test items with content standards is a critical source of
content-related validity evidence within high-stakes testing and accountability
frameworks. Typically, alignment studies are conducted with panels of experts
providing qualitative judgments on the degree of alignment with each item in the
test and the representative content standards, from which various summary
statistics are then calculated from these judgments [e.g., categorical
concurrence, balance of representation; @webb99]. In this paper, we propose an
approach that builds upon traditional methods for alignment but leverages the
use of text-based machine learning procedures, specifically topic modeling, to
identify text-based clusters within the content standards. The probability that
each item from a statewide assessment aligns with each cluster/topic can then be
estimated as an additional source of content-related validity evidence. We
discuss the utility of this approach, and show how visualizations can be used to
evaluate to overall coverage of the content standards by the test items.
class             : "man, fleqn, noextraspace"

header-includes:
    - \raggedbottom
    - \setlength{\parskip}{0pt}
    - \usepackage{xcolor}
output            : papaja::apa6_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE)

library(tidyverse)
library(rio)
library(here)
library(janitor)
library(tidytext)
library(topicmodels)
library(ldatuning)
library(scales)
library(patchwork)
#devtools::install_github("dgrtwo/drlib")
library(drlib)
#library(extrafont)
#font_import()

theme_set(theme_minimal(20) )
# +
#             theme(strip.text.x = element_text(family = "Times New Roman"),
#                   plot.title = element_text(family = "Times New Roman"),
#                   axis.text = element_text(color = "gray20",
#                                            face = "plain",
#                                            family = "Times New Roman"),
#                   axis.title = element_text(color = "gray20",
#                                             family = "Times New Roman",
#                                             face = "plain"),
#                   legend.text = element_text(family = "Times New Roman"),
#                   legend.title = element_text(family = "Times New Roman")))
```

```{r load_data}
standards <- import(here("data", "8gradescience.xlsx"), 
                    setclass = "tbl_df") %>% 
  clean_names()

webbwords <- import(here("data", "stopwords-webb.xlsx"),
                    setclass = "tbl_df")

items <- import(here("data", "G8_Sci_Items.xlsx"),
                setclass = "tbl_df") %>% 
  clean_names()
```

```{r doc_term_matrices}
# Create the document term matrix for standards
dtm <- standards %>% 
  select(code, ngss_standard) %>% 
  unnest_tokens(word, ngss_standard) %>% 
  anti_join(stop_words) %>% 
  anti_join(webbwords) %>% 
  filter(word != "explanation", # Filter out any additional words we don't want in there
         word != "results") %>% 
  group_by(code) %>% 
  count(word) %>% 
  ungroup() %>% 
  cast_dtm(code, word, n)

# Create a document term matrix for items 
idtm <- items %>% 
  select(item_id, standards, prompt, option_a, option_b, option_c) %>% 
  unite(text, prompt, option_a, option_b, option_c, sep = " ") %>%
  unnest_tokens(word, text) %>%
  filter(word != "a",
         word != "b",
         word != "c") %>%
  anti_join(stop_words) %>% 
  anti_join(webbwords) %>% 
  group_by(item_id) %>% 
  count(word) %>% 
  ungroup() %>% 
  cast_dtm(item_id, word, n)
```

For the last two decades, the landscape of education has evolved through the passage of legislation aimed at improving student outcomes through standards driven accountability such as the No Child Left Behind Act [@nclb02] and its more recent renewal as the Every Student Succeeds Act [see @mcguinn16]. This has led to a vast increase in standardized educational assessment use. As part of this movement, an intensive focus on alignment has also emerged due to the need for congruence between assessments, standards, curricula, and instruction to effectively impact student outcomes (e.g. [@porter02]). 

The goal of alignment is to “create a coherent educational system that conveys a clear and unified message about expectations and goals” [@vockley09, p.8]. Alignment studies are commonly conducted with standards-based assessments as a means of evaluating content-related validity evidence, with panels of experts judging the linkage between the content represented in the standards and the content represented in the test items [@sireci07; @webb97]. Content-related validity evidence is one of the five major sources of validity evidence outlined by the *Standards for Educational and Psychological Testing* [@standards14] and is a critical component of the "overall evaluative judgment" [@messick95, p. 741] of the validity of test scores for a given use. However, alignment studies are difficult to execute and design, even under ideal conditions. The chosen methodology, number of participants, professional judgments, consensus through discussion or averaged by ratings, costs associated with travel, and time are elements for consideration [see @anderson15]. Alignment studies are often time consuming and cost intensive [@anderson15]. 

According to Kane (2006) content validity is necessarily based on judgments, is more subject to confirmation bias, and lacks the objectivity of other sources of validity evidence, such as criterion-related evidence. This paper adds to the literature on content-related validity evidence through the use of text mining. Specifically, content-related validity evidence is provided by the extent to which test items and content standards display *textual congruence*; that is, to what extent does the text within the items align with the text used in the content standards? We use topic modeling, a text-based machine learning approach, to cluster groups of standards based on frequency of word co-occurrence. These clusters can then be evaluated for substantive meaning, and evaluated in relation to the standards themselves (i.e., whether the empirically-derived clusters match the standards domains or other standards groupings). The extent to which the text within individual test items corresponds with these clusters can be directly evaluated as an additional, alternative source of content-related validity evidence which can supplement, though not replace, information gained from alignment studies.

Topic modeling is a probabilistic method for identifying latent topics in text based documents. Its history stems from qualitative content analysis and latent semantic analysis [@mohr13]. Rather than the researcher predetermining the topics to be analyzed and coded, however, the topics emerge from corpora of text based on frequency of co-occurrence (i.e., text-based correlations). Topic modeling has advanced the field of text analysis from identifying specified words through a deductive approach where topics are pre-identified, to a more inductive approach where meaning is allowed to emerge through a corpus of texts [@mohr13]. Topic modeling is still relatively new in the scheme of text-based analytic research. Blei and colleagues [@blei03] introduced Latent Dirichlet Analysis (LDA) in 2003, which has become perhaps the most common topic model estimation procedure. Prior to LDA being introduced, inductive themes or latent meanings in text were achievable primarily through qualitative analysis [@nikolenko17]. Topic modeling, however, has promising potential in a broad array of contexts, including recent studies using LDA in combination with sentiment analysis for improved efficiency and accuracy when examining sentiments of consumer comments about products [@lin09], improved understanding of political themes across a range of documents[@hagen18], and the aggregation of results across scientific studies despite differences in terminology and fields of inquiry [@gefen17]. Topic modeling has even shown emerging potential to produce similar results to traditional qualitative frameworks, such as grounded theory, on qualitative survey data when utilizing a semi-supervised form of LDA [@nikolenko17]. The increasing accuracy of such techniques therefore provides support for potential use in expanded applications. 

In this paper we explore the use of topic modeling and specifically LDA as a source of content-related validity evidence to evaluate the textual congruence (alignment) between test items and content standards. We adopt a similar theoretical framework to alignment studies [@porter02; @webb97; @webb99], but from a text-based machine learning perspective, which may provide an additional triangulating source of alignment and content-related validity evidence. Further, given that the analyses themselves are relatively fast and less cost- and labor-intensive than alignment studies, topic-modeling may be a viable source of diagnostic information when developing test items.

# Method 
In this section, we describe the measures used for our specific application, as well as the standards they were designed to measure. We then discuss our analysis (topic model) in detail, including how we evaluated the number of topics to be extracted from the standards, and how the results can be used to evaluate the overall textual congruence between the content standards and the test items.

## Content Standards

Science instruction and assessment in Oregon are based on the *Next Generation Science Standards* (NGSS), which were adopted statewide in 2014. The NGSS are are based on the Framework for K-12 Science Education developed by the National Research Council (NRC) and are built around Performance Expectations, rather than more narrowly-defined standards of academic content knowledge, in each of grades K-5, and grade bands 6-8 (middle school) and 9-12 (high school). NGSS Performance Expectations are situated within one of four Domains (Physical [PS], Life [LS], Earth/Space [ESS] and Engineering Design [ED]) and organized into Arrangement Clusters (groups of interrelated Performance Expectations, e.g., Earth’s Place in the Universe [cluster ESS1]). Together, Performance Expectations unify three dimensions of science learning, requiring that students understand and apply *science and engineering practices* as they master *disciplinary core ideas* and engage *crosscutting concepts*, which span science (@nrc15). We evaluate the textual congruence between the the middle school (Grades 6-8) NGSS Performance Expectations, and the Grade 8 Alternate Assessment based on Alternate Achievement Standards (AA-AAS) in Oregon, as described more fully below.
<!-- 
Science Grade 8:
Domain	
LFS	= 9
PHS	= 7
ESS	= 6
ETS	= 2
TOTAL = 24
Note. LFS = Life Science Standards. PHS = Physical Sciences. ESS = Earth and Space Sciences. ETS = Engineering, Technology, and Applications.

The representation ratios (above) can be calculated by dividing the standards by the total within each respective column. For example, in Grade 8 Science, approximately 25% of the items are in the Earth and Space Sciences domain, as that domain has 6 written Essentialized Standards (EsSt) out of the total of 24 (6/24 = 25%).

We evaluate the textual congruence between the Grade 8 *Next Generation Science Standards* (NGSS) and the text used in the AA-AAS item stems and response options, as described in more detail below.
 -->

## Measures
Our application utilized the science portion of the Grade 8 statewide AA-AAS in Oregon, designed for students with the most significant cognitive disabilities [SWSCDs; @usdoe05]. The AA-AAS is an alternative assessment to the statewide general assessment, with up to 1% of students eligible for reporting purposes (note, there is a 1% reporting cap, not participation cap). SWSCDs are typically characterized by significantly below average general cognitive functioning. Commonly, this includes students with intelligence test scores two or more standard deviations below the mean on a standardized, individually administered intelligence test, occurring with commensurate deficits in adaptive behavior. Further, the cognitive disability must significantly impact the student's educational performance and ability to generalize learning from one setting to another. Students with the most significant cognitive disabilities in general, require highly specialized education and/or social, psychological, and medical services to access an educational program. These students may also rely on adults for personal care and have medical conditions that require physical/verbal supports, and assistive technology devices. These intensive and on-going supports and services are typically provided directly by educators and are delivered across all educational settings. Operationally, any student in Oregon with an individualized education program (IEP) is eligible to take the AA-AAS rather than the statewide general assessment, and the recommendation for the most appropriate assessment is guided by the IEP team.

Title 1 Federal Regulations published on December 9, 2003 [@usdoe03], prompted revisions to Oregon's AA-AAS to increase the  accessibility of items by reducing the depth, breadth, and complexity (RDBC) of the content within the test items. This essentialization process involved RDBC of the Common Core State Standards (CCSS), State Science Standards, and the Next Generation Science Standards (NGSS) in order to establish a performance expectation that was relevant and accessible for SWSCDs, while maintaining the highest possible standards of rigor. Complexity was reduced by: (a) focusing on essential content; (b) simplifying the process verb; and, (c) eliminating inappropriate delimeters. All essentialized standards were written at three levels of complexity: *Low*, *Medium*, and *High*.

## Analyses
Our process included (a) modeling and identifying key domains/topics within the statewide content standards based on frequency of word co-occurrence; (b) applying the model to new text, in the form of words represented in the test items through either the stem/prompt or the answer options; and (c) evaluating the overall mapping of items within the test to the modeled domains/topics, including the relative representativeness of each domain/topic determined through a combined approach of statistical analyses and expert judgement. We consider the evidence gained from these analyses as supplemental to formal alignment studies, and potentially useful as a diagnostic tool during test development.

### Topic modeling
Topic modeling is akin to exploratory factor analysis (EFA), where latent variables  (topics) are estimated based on the probability that the words within the topic will co-occur [see @mohr13]. Latent Dirichlet Allocation (LDA), introduced by @blei03, is perhaps the most common estimation procedure for topic models and was the approach used here. LDA is guided by the principles that "every document is a mixture of topics...[and] every topic is a mixture of words" [@silge17, p. 90]. This implies that each document may include words across a variety of topics, and, while the words within a given topic will generally be unique to the given topic, words can be shared between topics. LDA simultaneously estimates both the mixture of topics within a document and the mixture of words within a topic. The $\beta$ matrix then reports on the estimated probability that each word belongs to (or was generated by) a given topic, while the $\gamma$ matrix reports on the probability that each topic is represented within a given document. The ``documents" typically represent discretized instances of the overall corpus, such as a collection of blog posts or newspaper articles. In our application, we treat each NGSS Performance Expectation as a document and evaluate the representativeness of topics within each standard. For example, in a two-topic solution hypothetical Standard 1 may be composed of 25% Topic 1 and 75% Topic 2, while hypothetical Standard 2 is composed of 98% Topic 1 and 2% topic 2. The topics themselves are then represented by the corpora of words, with each word having a different modeled probability of having been generated by the corresponding topic. Post-hoc substantive labels are generally assigned to topics through expert evaluation of the top *n* words within each topic (typically 10-20), based on the $\beta$ matrix. 

As with EFA, perhaps the most difficult aspect of topic modeling is determining the number of topics (factors) to extract, which must be determined *a priori*. Models with different numbers of topics can provide different results and different conclusions about the underlying text. In our application, we relied on a combination of statistical evidence with expert judgment[^1]. From a statistical point of view, we relied upon four tests, as delineated by @arun10, @cao09, @deveaud14, and @griffiths04. Briefly, the method outlined by @arun10 is based on the KL-Divergence of two salient distributions, viewing LDA as a matrix factorization procedure, with the goal of minimizing this value. The @cao09 method relies on topic density through average cosine similarity (minimized), while the @deveaud14 method uses a similar approach but rely on the Jensen-Shannon distance between topic distributions (maximized). Finally, the method proposed by @griffiths04 uses Gibbs sampling with the posterior sampled such that the harmonic mean of the sampled log-likelihoods is maximized. See @houliu18 for a discussion on performance of these various indicators. We used these tests to evaluate the fit of models with 2 to 25 topics.


[^1]: It is important to note that we consider the topic model preliminary, and we welcome feedback from the field. All code used to estimate the models (and produce this manuscript) are publicly available at the following link: https://github.com/datalorax/text-analysis-content-validity

Following the evaluation of topics, we found a range of topics which appeared to reasonably minimize or maximize each of the criteria (as displayed in the results section). These topics were then reviewed for substantive interpretation and a final topic model was arrived upon through the evaluation of two separate science content experts. This model then represented our final trained model. The probability that each AA-AAS test item was represented by each topic was then estimated. The model was therefore trained on the words within the standards, and we can then evaluate whether the words used in the items correspond to those patterns.

As part of data preprocessing, we removed stop words (common words like "of", "a", "the", "and", "is", etc.) from the *onix*, *SMART*, and *snowball* lexicons [@lewis04; @onix; @snowball], as implemented in the *tidytext* R package [@silge16]. Additionally, we removed verbs associated with Webb's depth of knowledge levels [@webb02]. These  removals helped ensure the topics were clustered around content-related words, rather than specific verbs prevalent throughout the standards, or overly common words. Topics were estimated using the *textmodeling* package [@grun11], while the evaluation of the number of topics to extract was conducted with the *ldatuning* package [@nikita16], both of which are extensions to the  R statistical computing environment [@r]. Data were prepared using the *tidyverse* suite of packages [@wickham17], with all plots produced using the *ggplot2* package [@wickham16], and multiple plots assembled using the *patchwork* package [@pedersen17].

# Results
In this section, we first discuss the selection of the optimal number of topics. We then discuss the mapping of topics to standards, and words to topics. Finally, we present the results of the trained model to the items within the Grade 8 Science portion of the AA-AAS in Oregon, and specifically delineate the results by whether they were written to be of *low*, *medium*, or *high* difficulty. 

## Number of Topics
Figure 1 displays the optimal number of topics to be extracted across each of the four criteria. As would be expected, the different criteria suggested differing numbers of topics. Relying only on the criteria suggested by @arun10 would lead us to the conclusion that, essentially, the more topics estimated the better the fit, while the opposite conclusion would be reached if only evaluating the results relative to the criteria outlined by @deveaud14. The @cao09 and @griffiths04 criteria were more nuanced and suggested differing, but similar ranges of topics. After six topics, however, a marked increase in the @cao09 criteria was observed, indicating a poorer fit, while only a moderate increase (indicating a better fit) was observed with the @griffiths04 criteria. Taken together, these results suggested that between three [the minimum value for the @cao09 criteria] and six topics should be extracted (as displayed by the shaded rectangle in the background of Figure 1). Each of these topic solutions were therefore evaluated by expert judgement for substantive meaning.

```{r n-topics, fig.cap = "Optimal Topic Selection. Optimal number of topics displayed according to four separate criteria. Shaded rectangle displayed the range in which topics were evaluated for substantive meaning. ", fig.width = 6.5, fig.height = 8}

search_standards <- FindTopicsNumber(dtm, 
                                     topics = 2:25,
                                     metrics = c("Griffiths2004", 
                                                 "CaoJuan2009", 
                                                 "Arun2010", 
                                                 "Deveaud2014"),
                                     method = "Gibbs",
                                     control = list(seed = 77),
                                     mc.cores = 4L)

stats_standards <- apply(search_standards[ ,-1], 
      2, 
      function(x) rescale(x, c(0, 1), range(x))) %>%
  as_tibble() %>%
  rowid_to_column("Topics") %>%
  gather(metric, val, -Topics) %>%
  mutate(criterion = ifelse(metric == "CaoJuan2009" |
                            metric == "Arun2010",
                            "min",
                            "max"),
        metric = case_when(metric == "Griffiths2004" ~ 
                                        "Griffiths & Steyvers, 2004", 
                           metric == "CaoJuan2009" ~ 
                                        "Cao et al., 2009", 
                           metric == "Arun2010" ~
                                        "Arun et al., 2010", 
                           metric == "Deveaud2014" ~
                                        "Deveaud et al., 2014"))

p1 <- ggplot(filter(stats_standards, criterion == "min"),
       aes(Topics, val)) +
  annotate("rect",
           xmin = 3, xmax = 6,
           ymin = 0, ymax = 1,
           fill = "magenta",
           alpha = 0.2) +
  geom_line(lwd = 1.2, 
            color = "cornflowerblue") +
  geom_point(color = "gray40",
             size = 3) +
  facet_wrap(~metric) +
  labs(x = NULL,
       y = NULL,
       title = "Minimize") +
  theme(plot.title = element_text(hjust = 0.5,
                                  vjust = -3,
                                  face = "bold"),
        panel.spacing = unit(2, "lines"))


p2 <- ggplot(filter(stats_standards, criterion == "max"),
       aes(Topics, val)) +
  annotate("rect",
           xmin = 3, xmax = 6,
           ymin = 0, ymax = 1,
           fill = "magenta",
           alpha = 0.2) +
  geom_line(lwd = 1.2, 
            color = "cornflowerblue") +
  geom_point(color = "gray40",
             size = 3) +
  facet_wrap(~metric) +
  labs(x = "Number of Topics",
       y = NULL,
       title = "Maximize") +
  theme(plot.title = element_text(hjust = 0.5,
                                  vjust = -3,
                                  face = "bold"),
        panel.spacing = unit(2, "lines"))

p1 / p2
```

Two science content experts with strong familiarity with the NGSS evaluated the substantiveness of three through six topic solutions, independently, resolving differences in naming the topics through post-hoc collaboration. In general, the content experts began with a three-topic solution and first considered each word within a given extracted topic individually, including their acceptation and known use within the NGSS. Second, the two experts considered their combined meaning within each extracted topic as being a representation of a possible construct (or topic of study) in the broad field of science, while noting the value of $\beta$ for each word, and gave each topic a substantive scientific name (see Figure 3). Last, the experts examined each named topic, representing one or more scientific constructs based on the number of extracted topic solutions, for independence, overlap, and stability across solutions. For example, once the experts examined each individual and combination of words and then applied a substantive name to each topic within the three-topic solution set, they went through the same process for a four-topic solution, and then compared the substantive names between the two solutions for independence, overlap, and stability (i.e., did each topic in the lower solution remain, were new topics created that were independent from others and that had substantive scientific meaning). Independently, the content experts settled on the five-topic solution as best representing the data. Specifically, increasing the number of topics from three to four and from four to five led to the addition of substantively new and independent topics with little overlap. However, adding a sixth topic led to substantive overlap with at least two other extracted topics, and thus, the five-topic solution was deemed most appropriate. Table 1 displays substantive labels collaboratively determined by the two content experts for each of the topics derived from our final (five-topic) model.

```{r, results = "asis"}
tibble(Topic = 1:5,
       `Substantive Label` = 
    c("Analyzing data and using evidence to understand organisms and systems",
      "Using scientific evidence to understand Earth systems",
      "Energy",
      "Genetic information",
      "Scientific solutions")) %>%
papaja::apa_table(caption = 
             "Substantive Labels Assigned to Final Topic Solution")
```

## Mapping Topics to Standards and Words to Topics


```{r finalized-model}
tm_raw <- LDA(dtm, k = 5, control = list(seed = 1234))

betas <- tidy(tm_raw, "beta") 
gammas <- tidy(tm_raw, "gamma") 

ess1_1 <- filter(gammas, document == "S08ESS1.1")
ess2_4 <- filter(gammas, document == "S08ESS2.4")

```

Figure 2 displays a heatmap of the $\gamma$ values across topics for each of the 59 performance expectations evaluated. Brighter colors represent a higher likelihood of the given topic being represented by the given standard. For example, Performance Expectation ESS1.1 (bottom) is represented almost entirely (`r paste0(round(ess1_1$gamma[1], 2)*100, "%")`) by Topic 1: *Analyzing data and using evidence to understand organisms and systems* . Performance Expectation ESS2.4, however, is represented partially (`r paste0(round(ess2_4$gamma[1], 2)*100, "%")`) by Topic 1, and partially (`r paste0(round(ess2_4$gamma[2], 2)*100, "%")`) by Topic 2: *Using scientific evidence to understand Earth systems*. These heatmaps can assist in deriving substantive meaning from the topics, given that the standards that predominately make up a topic can be investigated. Once substantive meaning is assigned, however, the topics can likewise help bring new meaning back to the standards, as themes may emerge that were otherwise not apparent. 


```{r heatmap, fig.cap = "Heatmap of Gamma Values. Cells displayed in brighter colors have higher gamma values, representing a greater likelihood that the given topic is represented by the given standard.", fig.height = 7.5, fig.width = 6.5}
gammas_max <- gammas %>%
  group_by(document) %>%
  filter(gamma == max(gamma)) %>%
  rename(assigned = topic) %>%
  select(-gamma)

gammas <- left_join(gammas, gammas_max)

ggplot(gammas, aes(topic, fct_reorder(document, assigned))) +
  geom_tile(aes(fill = gamma),
                color = "gray40") +
  scale_fill_viridis_c(name = "\nGamma ", 
                       option = "magma",
                       guide = guide_colorbar(
                                  direction = "horizontal",
                                  label.position = "top",
                                  ticks = FALSE,
                                  breaks = seq(0, 1, 0.25),
                                  barwidth = grid::unit(3.5, "in"),
                                  barheight = grid::unit(0.2, "in")),
                       limits = c(0, 1),
                       begin = 0, 
                       end = 1) +
  scale_x_continuous(breaks = 1:5, labels = 1:5) +
  labs(x = "Topic", 
       y = "Standard Code") +
  coord_cartesian(expand = FALSE) +
  theme(plot.margin = margin(1.4, 0.2, 0.2, 0.2, "cm"),
        legend.position = c(0.62, 1.05),
        legend.title=element_text(size = 12,
                                  color = "gray40"),
        legend.text=element_text(size = 8,
                                 color = "gray40"),
        axis.text.y = element_text(size = 7),
        legend.key.width = unit(1.2, "cm"))

```


The top 15 words within each topic are displayed in Figure 3, according to their estimated $\beta$ values. Note that in many instances there were ties among beta values around 15, those selected for display were chosen randomly. For example, if if the 13^th^ to 18^th^ words all had the same $\beta$ value, only the 13^th^ to 15^th^ would be displayed. Each topic has been labeled according to its identified substantive label.


```{r word-freq, fig.cap = "Beta Values. Highest probability of words generated by topic. Note the substantive labels were assigned post-hoc.", fig.width = 12, fig.height = 12}
pd <- betas %>%
  group_by(topic) %>%
  arrange(desc(beta)) %>%
  slice(1:15) %>%
  ungroup() %>%
  mutate(topic = factor(topic, 
                            levels = 1:5,
                            labels = c("Organisms & Systems",
                                       "Earth Systems",
                                       "Energy",
                                       "Genetic Information",
                                       "Scientific Solutions")),
         reordered = reorder_within(term, beta, topic))

ggplot(pd, aes(reordered, beta)) +
  geom_col(fill = "#6272F8",
           alpha = 0.7) +
  coord_flip(clip = "off") +
  scale_y_continuous(breaks = seq(0, 0.09, 0.03),
                     labels = seq(0, 0.09, 0.03)) +
  scale_x_reordered(name = "") +
  facet_wrap(~topic, scales = "free_y") 

# Need to fix this so it shows up as "earth's" instead of "earth..s"

```

## Predicting Items to Topics

In addition to the topic-model providing information about the interrelated nature of the NGSS performance expectations, the finalized model can also be used for predictive purposes. That is, new text can be provided to the model, and topic-level predictions can be made. Specifically, each word from the next text is assigned a probability that it was generated from each topic. We used this approach with text from Grade 8 science items from the AA-AAS used in Oregon. Importantly, all text from a given item, including the item prompt and the item options, were used when making predictions. Prediction is a potentially valuable attribute in the alignment process because it can be utilized to not only predict probablities of text used but also evaluate content coverage and representativeness within the assessement.

```{r radar-plot-data}
# Predict the topic for each item
posteriors <- posterior(tm_raw, newdata = idtm)

coord_radar <- function (theta = "x", start = 0, direction = 1) {
 theta <- match.arg(theta, c("x", "y"))
 r <- if (theta == "x") 
        "y"
      else "x"
 ggproto("CoordRadar", CoordPolar, theta = theta, r = r, start = start, 
      direction = sign(direction),
      clip = "off",
      is_linear = function(coord) TRUE)
}

probs <- posteriors$topics %>% 
  as.data.frame() %>% 
  mutate(item = rownames(.)) %>% 
  tbl_df() %>% 
  gather(topic, probability, -item) %>% 
  mutate(Level = str_extract(item, "L|M|H"),
         Level = factor(Level, 
                        levels = c("L", "M", "H"),
                        labels = c("Low", "Medium", "High")),
         topic = factor(topic,
                        levels = 1:5,
                        labels = c("Organisms & Systems",
                                   "Earth Systems",
                                   "Energy",
                                   "Genetic Information",
                                   "Scientific Solutions")))

```

Figure 4 displays the probability of the text used within a random sample of nine items being generated from each of the five modeled topics. Random items 2, 5, and 7 all did not include any text that could be classified by our model, and the probability was equally distributed across the five topics. Note that this was expected, and is likely to be a more commonly observed outcome when working with data from AA-AAS assessment systems, given the RDBC process discussed above, and that textual density within items is generally  minimized by design. The equal spread of probability across topics is therefore not an indication that the items does not align with the content standards, but rather that the text represented in  the item (if any) was not represented within our model. Random Items 4, 6, and 8 all clearly aligned with a single topic, while random items 1, 3, and 9 had their probability split between two topics. Mapping these topics back to the standards they compose can therefore provide additional information about the items and their linkage to the content standards.

```{r, fig.width = 6.5, fig.height = 8, fig.cap = "Probability of topics by item. Random sample of nine items displayed."}
set.seed(3)
samp <- probs %>% 
  sample_n(9)

probs %>% 
  filter(item %in% samp$item) %>% 
  mutate(item = as.numeric(as.factor(item)),
         item = factor(item,
                       levels = 1:9,
                       labels = paste0("Random Item ", 1:9))) %>% 
ggplot(aes(topic, probability, color = item)) +
  geom_polygon(aes(group = Level, fill = item, color = item), alpha = 0.7,
               lwd = 1.3) +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_discrete(breaks = c("Organisms & Systems",
                              "Earth Systems",
                              "Energy",
                              "Genetic Information",
                              "Scientific Solutions"),
                   labels = str_wrap(c("Organisms & Systems",
                                       "Earth Systems",
                                       "Energy",
                                       "Genetic Information",
                                       "Scientific Solutions"), 
                                     width = 12)) +
  facet_wrap(~item) +
  coord_radar() +
  guides(fill = "none",
         color = "none") +
  labs(x = "",
       y = "") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 7),
        panel.spacing.x = unit(2, "cm"),
        panel.spacing.y = unit(0, "cm"),
        strip.text = element_text(size=10))
```

The model-based predictions from the text used within the items can also be used to evaluate the content coverage. In our application, test items were theoretically designed to be of *Low*, *Medium*, and *High*  difficulty, as discussed in the methods. We were therefore interested in if the coverage of the topics was equal across each type of item. Figure 5 displays a summary of the overall topic coverage by item type. Essentially, the average probability of items representing each of the five topics is displayed on the log scale, specifically  $log(p(x_i + 1))$. The thick gray band represents the expected probability, if all topics were equally represented. Items in the low category, for  example, were estimated as slightly under-representing all topics, with the exception of *Organisms and Systems*, which was highly over-represented. This same pattern was present for the items written to be of *Medium* difficulty, although it was even more severe. For the *High* items, the *Genetic Information* topic was underrepresented, but overall the coverage looks considerably better. Although the evidence is not definitive, this type of information could potentially be used to guide subsequent investigations of content representativeness. 

```{r, fig.width = 6.5, fig.height = 12, fig.cap="Overall Content Coverage"}
ref <- tibble(x = c(1:5, 1.05), y = rep(0, 6))

probs %>% 
  group_by(topic, Level) %>% 
  summarize(n = n(),
            prob = sum(probability))  %>% 
  mutate(prob = log((prob/n)*5)) %>% 
  ungroup() %>% 
ggplot(aes(x = topic, y = prob)) + 
  geom_polygon(aes(group = Level, fill = Level, color = Level), alpha = 0.7) +
  geom_path(aes(x, y), ref, color = "gray40", lwd = 1.2) +
  scale_color_brewer(palette = "Dark2") +
  scale_fill_brewer(palette = "Dark2") +
  scale_x_discrete(breaks = c("Organisms & Systems",
                              "Earth Systems",
                              "Energy",
                              "Genetic Information",
                              "Scientific Solutions"),
                   labels = str_wrap(c("Organisms & Systems",
                                       "Earth Systems",
                                       "Energy",
                                       "Genetic Information",
                                       "Scientific Solutions"), 
                                     width = 12)) +
  facet_wrap(~Level) +
  coord_radar() +
  guides(fill = "none",
         color = "none") +
  labs(x = "",
       y = "") +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 7),
        panel.spacing.x = unit(1, "cm"),
        panel.spacing.y = unit(0, "cm"))
```

# Discussion

Content validity is critical to the overall evaluative judgment of the validity of a test for a given use [@kane06; messick95] and a is core element examined in alignment studies. This paper introduced text mining procedures, specifically topic modeling [@mohr13], to evaluate the textual congruence between content standards test items. This new approach demonstrates potential to supplement the methods of current alignment studies by providing a triangulated source of content validity data that includes elements of objectivity and cost-effectiveness. In addition, this approach holds potential as a means of learning more about the content standards themselves (i.e., identifying groups of standards through previously unobserved themes), test items, and the textual link between test items and standards as part of the iterative assessment development process. 

Specifically, information from topic modeling could be used to supplement data collected through formal alignment studies by triangulating evidence of the alignment for individual items. If the same information is provided by sources, it may help reinforce the overall substantive conclusions about particular items or a test as a whole. If the evidence does not agree, however, then it may serve as a prompt to collect more data and conduct further investigations. In addition, the analyses could potentially inform item and test development *during* the developmental process. That is, the analysis could serve as a diagnostic tool to better understand the content coverage of a particular test and when key vocabulary may be missing from a particular item. From a cost-benefit perspective, it is much cheaper to conduct an analyses of data in-house than to conduct formal alignment studies. Part of the benefit of an analytic approach is that the analyses could be conducted much more regularly to inform the iterative test documentation/validation process. 

## Limitations and Future Directions
There are several limitation to the approach discussed in this paper that should be kept in mind. First, the results depend highly upon the chosen topic model, including both the number of topics and the substantive meaning provided to the topics. We view the topic model used here as preliminary, and in need of further external validation. We have therefore published figures similar to Figures 2 and 3 for topic solutions between 2 to 16 topics within the GitHub repository that houses all of our analyses[^2]. Ideally, the topic model itself could be refined over time, perhaps through a semi-supervised approach to topic modeling [@lu08], that would represent expert consensus across the field. By making all of our work public, we hope to encourage collaboration and potentially "crowd source" an optimal model - an approach that has been successfully applied in other fields [e.g., @arganda15; bentzien13]. Once the model reached a stable point, an application programming interface (API) could even be constructed such that test developers could submit the text from their items and obtain immediate feedback, perhaps in a similar for to the results provided here.

[^2]: See here: https://github.com/datalorax/text-analysis-content-validity

It is also the case that some items, by design, have little to no text. In this case, our model essentially fails, because it cannot relate the item to any topic, and the probability is equally dispersed across the topics. We view these cases as potential for further review. In other words, the item may align well with the content standards and contribute meaningfully toward the content-related validity evidence of the test, but not include any text. A more long-term solution would be to utilize other areas of machine learning, such as image recognition. This would include first creating a dataset with items with no text, and providing the standard(s) with which the item was aligned. The algorithm could then begin to learn how different image features relate to item alignment. Again, however, this would require the algorithm being iteratively trained, with different algorithms being needed for different content areas. Expert consensus would again need to be established. However, if text mining could be used in combination with image recognition (and one could potentially train the image recognition software to learn the topics, so the algorithms would work together), a more powerful content-related validity evidence system could be established. We again, however, would view such a system as diagnostic and a source of triangulating evidence, with human judgment ultimately being the "gold standard". 

Finally, our specific application utilized the NGSS content standards and a science test. In some ways, science is "easy" to evaluate linkage between tests and items, because key vocabulary must be represented in each. It therefore remains to be seen how this approach would work with other content areas. Reading would seem to follow rather naturally; for example if a standard discussed verbs, an item may as students to identify the verb within a sentence. Because the word "verb" would occur in both, the match would be found. However, there are other instances where the mapping between the text used in content standards and items may be disparate, but yet aligned. Mathematics may be particularly tricky but, again, more research would be needed. For example, the mathematical symbols themselves could be treated as "words". For topics such as geometry, the image recognition approach discussed above may be more appropriate. 

## Conclusions
Machine learning is rapidly changing many fields. In this paper, we introduced a text-mining procedure that could provide additional evidence and context to alignment studies, and potentially serve as a diagnostic tool during item development. If such a method were to be applied in operational use, however, it would be critical to communicate the shortcomings of the approach with stakeholders, and that the results not be relied upon too heavily. Attaching numbers and probabilities to words and items may communicate to stakeholders a more "objective" approach, but as discussed above, this is not necessarily the case. As a general diagnostic tool, and as a supplemental source of evidence, however, we believe the approach discussed here holds considerable promise. The practical costs associated with alignment studies are high. Ideally, the alignment study should essentially provide confirmatory evidence, while ad-hoc text analyses could serve as a more diagnostic and exploratory tool. 

The gathering of content-related validity evidence is dominated by a single method - alignment studies. Machine learning approaches may help to broaden the evidential base. In some cases, specifically in the early grades (pre-K; K-2) tests may not have content standards, and the approach discussed here may seem infeasible. However, formal alignment studies are also challenging if not impossible at these levels and, while predictions may not be able to be made toward topics represented in content standards, much could be learned by mining the items themselves and evaluating the topics therein. In sum, while the work presented here is preliminary, we feel text mining and topic modeling as a whole hold considerable promise for the field of measurement and, from our perspective, have thus far been somewhat underutilized. We are hoping this approach will spur new discussion and innovation in the area of content validity, measurement development, and alignment studies given the potential added benefits realated to cost-effectiveness, objectivity, prediciton, and formative use in the iterative measurement development process. 

\newpage
# References
\begingroup
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}

<div id = "refs"></div>
\endgroup



